---
title: "BMI 585 - Final Project"
author: "Hope Mumme"
date: "12/9/2021"
output: html_document
---

This Markdown file contains example uses for all the functions in the package bmi585hmumme. You can download the package with the following command in the R Console. Uncomment the following if you have not installed the package to your system. 

```{r Install}
#devtools::install_github("hmumme/bmi585hmumme")
```

Load the package and set up the Markdown file:

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(bmi585hmumme)
```

The following functions are broken down by section; each section represents the assignment the function came from. These examples can also be found in the help documentation for each function; type ?functionName to retrieve this. 


### newFunctions.R

##### unscale(x)

This function takes a numerical object from scale(), reverses the scaling and centering done (if any), and returns the un-scaled numerical object.

```{r unscale}
xS = scale(matrix(1:10, ncol = 2)) # scale a matrix
unscale(x = xS) # un-scale to get the original matrix
```

##### pcApprox(x,npc)

This function approximates x feature data using npc number of principle components.

```{r pcApprox}
x = matrix(c(4,5,2,10,2,8,4,5,6), ncol = 3) # feature data matrix
pcApprox(x, npc = 1) # approximate using 1 principle component
```

##### pcLollipop(x)

This function creates lollipop plots of all the principal component loadings in feature data x (must be numerical data frame or tibble). Each principle component will have its own plot of values for each variable in the data x. 

```{r pcLollipop}
x = data.frame(data = matrix(c(4,5,2,10,2,8,4,5,6), ncol = 3)) # feature data frame
pcLollipop(x) # create lollipop plots
```

### quiz02Functions.R

##### sensitivity(pred,truth)

This function calculates the sensitivity from predicted and true vectors. These vectors can either be 0s (negatives) and 1s (positives) or boolean. 

```{r sensitivity}
sensitivity(c(0,1,0),c(1,1,1)) # calculates sensitivity
```

##### specificity(pred,truth)

This function calculates the specificity from predicted and true vectors. These vectors can either be 0s (negatives) and 1s (positives) or boolean. 

```{r specificity}
specificity(c(0,1,1),c(0,0,1)) # calculates specificity
```

##### accuracy(pred,truth)

This function calculates the accuracy from predicted and true vectors. These vectors can either be 0s (negatives) and 1s (positives) or boolean. 

```{r accuracy}
accuracy(c(0,1,1),c(1,1,1)) # calculates accuracy
```

##### ppv(pred,truth)

This function calculates the PPV from predicted and true vectors. These vectors can either be 0s (negatives) and 1s (positives) or boolean. 

```{r ppv}
ppv(c(1,1,0),c(0,1,1)) # calculates PPV
```

##### f1(pred,truth)

This function calculates the F1 score from predicted and true vectors. These vectors can either be 0s (negatives) and 1s (positives) or boolean. 

```{r f1}
f1(c(1,1,1),c(1,1,0)) # calculates f1 score
```

### hw04Functions.R

##### boxMuller(n)

This function use the Box-Muller transformation to generate 2 vectors of random, normally distributed n values.

```{r boxMuller}
boxMuller(5) # generates two random, normally distributed vectors with 5 values each
```

### hw06Functions.R

##### twoSidedT(t,n)

This function calculates the p-value from a two-sided t-test statistic and the number of samples n.

```{r twoSidedT}
twoSidedT(3, 5) # calculates p-value
```

##### twoSizedZ(z)

This function calculates the p-value from a two-sided z score.

```{r twoSidedZ}
twoSidedZ(3) # calculates p-value
```

##### effectSize(x,g)

This function calculates the effect size from vector of values x and grouping variable g.

```{r effectSize}
effectSize(c(1.42, 8.10, 9.22, 20.92, 5.67, 0.45), c(0,0,1,1,0)) # calculates effect size
```

##### welchT(x,y)

This function performs a Welch's T test to compare the means of vector x and vector y. It will return a vector of length 5 with Welch's T statistic, degrees of freedom, p-value, mean of group x, and mean of group y.

```{r welchT}
welchT(c(1.42, 8.10, 9.22, 20.92, 5.67, 0.45), c(1.99, 8.09, 9.13, 9.10, 5.55, 0.11)) # calculate welch t-test statistics
```

### hw07Functions.R

##### minimumN(d)

This function finds the minimum number of samples needed for effect size d and power = 0.8, returns number of samples rounded up to nearest integer.

```{r minimumN}
minimumN(1.3) # calculate number of samples needed
```

##### chiSquareCounts(tib,x,y)

This function performs a chi square test of homogeneity between data in columns x and y of tib. Returns a p-value.

```{r chiSquareCounts}
data = dplyr::tibble("sex" = c(1,0,1),"group" = c("A","B","C"), "age" = c(40,15,33), "height" = c(63, 70, 68)) # create data
chiSquareCounts(data, "age", "height") # perform test
```

##### postHocPower(d,n1,n2)

This function estimates post hoc power from effect size and number of samples using 1000 simulations. 

```{r postHocPower}
postHocPower(d = 0.5, n1 = 20, n2 = 22) # estimate power
```

### hw08Functions.R

##### bhAdjust(p)

This function tests for significant p-values using Bonferroni-Holm adjustment. Returns logical vector showing which p-values are statistically significant after the BH adjustment. 

```{r bhAdjust}
bhAdjust(c(0.05, 0.0005, 0.01, 0.0225, 0.025)) # find which p-values are significant after BH adjustment
```

##### fdrAdjust(p)

This function tests for significant p-values using FDR adjustment. Returns logical vector showing which p-values are statistically significant after the FDR adjustment. 

```{r fdrAdjust}
fdrAdjust(c(0.05, 0.0005, 0.01, 0.0225, 0.025)) # find which p-values are significant after fdr adjustment
```

### hw11Functions.R

##### r2(pred,truth) 

This function calculates the R-Squared value from two numerical vectors of predicted/true values.

```{r r2}
yTrue = c(0,1,1,1) # true values
r2(pred = c(0,1,0.7,1), truth = yTrue) # find R^2
```

##### adjR2(pred,truth)

This function calculates the adjusted R-Squared value from two numerical vectors of predicted/true values and the number of predictors used d.

```{r adjR2}
yTrue = c(0,1,1,1) # true values
adjR2(pred = c(0,1,0.7,1), truth = yTrue, d = 2) # find adjusted R^2
```

